IMPC Mouse data - Variance in sex differences

### 2020-04-20 Prep & old/FZ 10-12-2019.Rmd
Add files via upload

# Apply some standard cleaning to the data
clean_raw_data <- function(mydata) {
  mydata %>%

@@

    # sort
    arrange(production_center, biological_sample_id, age_in_days) %>%

    # filter to groups with > 1 centre  
    group_by(parameter_name) %>%
    mutate(center_per_trait = length(unique(production_center, na.rm = TRUE))) %>%
    ungroup() %>% 
    filter(center_per_trait >= 2) %>%

@@

    # add grouping variable: these were decided based on functional groups and procedures 
    left_join(by="parameter_name", 
      read_csv(here("data", "ParameterGrouping.csv")) %>%
        select(-id)
  ) %>% 
  #### Assign unique IDs (per trait)
# each unique parameter_name (=trait,use trait variable) gets a unique number ('id')

# We add a new variable, where redundant traits are combined
#[note however, at this stage the dataset still contains nonsensical traits, i.e. traits that may not contain any information on variance]
  mutate(id = match(parameter_name, unique(parameter_name)))
}

# Load raw data - save cleaned dataset as RDS for reuse

data_raw <- load_raw(here("data","dr7.0_all_control_data.csv.gz"))

dir.create("export", F, F)

data_raw %>% 
  clean_raw_data() %>% 
  saveRDS("export/data_clean.rds")
```

For analysis we load the RDS created above:

```{r load}
data <- readRDS(here("export", "data_clean.rds")) 
```

Check length of different variables:

```{r}
length(unique(data$parameter_name)) # 232
length(unique(data$parameter_group)) # 161
length(unique(data$procedure_name)) # 26

@@

# Meta analysis

## Phase 1: across all traits

@@

data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min=0, age_center=100) {
  mydata %>%
    filter(
      age_in_days >= age_min,
      id == parameter
    ) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff) %>% 
    filter(!duplicated(biological_sample_id))
}
```

@@

for (t in 1:n) {
  tryCatch(
    {
      results <- data %>% 
        data_subset_parameterid_individual_by_age(t) %>%
        calculate_population_stats() %>%
        create_meta_analysis_effect_sizes()

@@

      results_alltraits_grouping2[t, 2:15] <- 
        c(f(cvr), f(cv), f(means), means$k, unique(results$trait))
    },
    error = function(e) {
      cat("ERROR :", conditionMessage(e), "\n")
    }
  )
}
```

Now that we have a "results" table with each of the meta-analytic means for all effect sizes of interest, we can use this table as part of the Shiny App, which will then be able to back calculate the percentage differences between males and females for mean, variance and coefficient of variance. We'll export and use this in the Shiny App. **Note that I have not dealt with convergence issues in some of these models, and so, this will need to be done down the road**

## Note Susi 31/7/2019: This dataset contains dublicated values, plus no info on what the "traits" mean. I will change Dan N's to one further belwo, that have been cleaned up already
# FILE TO USE: METACOMBO (around line 500)

# trait_meta_results <- write.csv(results_alltraits_grouping, file = "export/trait_meta_results.csv")

#### Merging datasets & removal of non-converged traits

@@

```{r}
procedures <- read.csv(here("export", "procedures.csv"))

results_alltraits_grouping$parameter_group <- data$parameter_group[match(results_alltraits_grouping$id, data$id)]
results_alltraits_grouping$procedure <- data$procedure_name[match(results_alltraits_grouping$id, data$id)]

results_alltraits_grouping$GroupingTerm <- procedures$GroupingTerm[match(results_alltraits_grouping$procedure, procedures$procedure)]
results_alltraits_grouping$parameter_name <- data$parameter_name[match(results_alltraits_grouping$id, data$id)]

meta1 <- results_alltraits_grouping
n <- length(unique(meta1$parameter_name)) # 232
```

@@

```{r}
meta_clean <- meta1[ !(meta1$id %in% c(84, 144, 158, 160, 161, 162, 163, 165, 166, 167, 168, 221, 222, 231)), ]
removed <- length(unique(meta_clean$parameter_name)) # 218

meta_problem <- results_alltraits_grouping[ (results_alltraits_grouping$id %in% c(84, 144, 158, 160, 161, 162, 163, 165, 166, 167, 168, 221, 222, 231)), ]
```

@@

meta1 <- meta_clean
# length(unique(meta1$procedure)) #18
# length(unique(meta1$GroupingTerm)) #9
# length(unique(meta1$parameter_group)) # 148 levels. To be used as grouping factor for meta-meta analysis / collapsing down based on things that are classified identically in "parameter_group" but have different "parameter_name"
# length(unique(meta1$parameter_name)) #218
```

@@

kable(cbind(meta1 %>% count(parameter_group))) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

meta1b <-
  meta1 %>%
  group_by(parameter_group) %>%
  summarize(par_group_size = length(unique(parameter_name, na.rm = TRUE)))
# this gives a summary of number of parameter names in each parameter group, now it neeeds to get merged it back together

@@

meta1_sub <- subset(meta1, par_group_size > 1) # 90 observations
meta1_sub$sampleSize <- as.numeric(meta1_sub$sampleSize)
```

```{r}
# nesting
n_count <- meta1_sub %>%
  group_by(parameter_group) %>%
  mutate(raw_N = sum(sampleSize)) %>%
  nest() %>%
  ungroup()

@@

```{r}
metacombo_final <- metacombo %>%
  group_by(GroupingTerm) %>%
  nest()

@@

# For all grouping terms
metacombo_final_all <- metacombo %>%
  nest()

@@

>>>>

# Apply some standard cleaning to the data
clean_raw_data <- function(mydata) {

  group <- read_csv(here("data", "ParameterGrouping.csv"))

  tmp <- 
    mydata %>%

@@

    # sort
    arrange(production_center, biological_sample_id, age_in_days)

    # filter to groups with > 1 centre  
    # **Review** - retaining merge here to retain ordering of data used in previous version
    # usually i would use mutate instead of creating sperate dataset
    merge(tmp, 
          tmp %>% group_by(parameter_name) %>%
    summarise(center_per_trait = length(unique(production_center, na.rm = TRUE)))
        )%>%
    filter(center_per_trait >= 2) %>%

@@

    # add grouping variable: these were decided based on functional groups and procedures 
    # **Review** - retaining match here to retain ordering of data used in rpevious version
    mutate(parameter_group = group$parameter_group[match(parameter_name, group$parameter_name)] ) %>%
    # usually i would use left_join
        # left_join(by="parameter_name", 
    #   read_csv(here("data", "ParameterGrouping.csv")) %>%
    #     select(-id)
    # ) %>%

    # Assign unique IDs (per trait)
    # each unique parameter_name (=trait,use trait variable) gets a unique number ('id')

    # We add a new variable, where redundant traits are combined
    #[note however, at this stage the dataset still contains nonsensical traits, i.e. traits that may not contain any information on variance]
    mutate(id = match(parameter_name, unique(parameter_name))) %>% 
    as_tibble()
}

# Load raw data - save cleaned dataset as RDS for reuse
data_raw <- load_raw(here("data","dr7.0_all_control_data.csv.gz"))
dir.create("export", F, F)

data <- data_raw %>% 
  clean_raw_data() 
saveRDS(data, "export/data_clean.rds")
```

For analysis we load the RDS created above and other datasets:

```{r load}
data <- readRDS(here("export", "data_clean.rds"))

procedures <- read_csv(here("data", "procedures.csv"))

@@

Check length of different variables / sample sizes:

```{r}
length(unique(data$parameter_name)) # 232 traits
length(unique(data$parameter_group)) # 161 parameter groups
length(unique(data$procedure_name)) # 26 procedure groups
length(unique(data$biological_sample_id)) # 27147 individial mice   #FZ added 11-12-2019
#number of males and females per strain per production center #FZ added 11-12-2019
sampleSizes <-data %>% group_by(production_center, strain_name) %>% count(biological_sample_id, sex) %>% count(sex) %>% print(n = Inf)
sampleSizes <- write.csv(sampleSizes, file = "export/sampleSizes.csv")
length(unique(sampleSizes$strain_name)) #9
length(unique(data$strain_name)) #9
```

# Phase 1: across all traits

@@

data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min=0, age_center=100) {
  tmp <- mydata %>%
    filter(
      age_in_days >= age_min,
      id == parameter
    ) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)# %>% 
#    filter(!duplicated(biological_sample_id))

  # still some individuals with multiple records (because same individual appear under different procedures, so filter to one record)
  j <- match(unique(tmp$biological_sample_id), tmp$biological_sample_id)
  tmp[j, ] 
  }
```

@@

for (t in 160:160) {
  tryCatch(
    {
      results <- data %>% 
        data_subset_parameterid_individual_by_age(t) %>%
        calculate_population_stats() %>%
        create_meta_analysis_effect_sizes()

@@

      results_alltraits_grouping[t, 2:14] <- c(f(cvr), f(cv), f(means), means$k)
      results_alltraits_grouping[t, 15] <- unique(results$trait)
    },
    error = function(e) {
      cat("ERROR :", t, conditionMessage(e), "\n")
    }
  )
}
```

--------
**REVIEW**: The above loop gives some errors, can these be fixed?
**FZ**: In the above function, we use 'tryCatch' and 'conditionMessage' to prevent the loop from aborting when the first error at row 84 is produced.
**FZ**  Convergence in the two listed non-converging cases can't be achieved by sensibly tweaking (other optim etc.).
**FZ**  Since we only learn about non-convergence in the loop, it's difficult to exclude the two traits beforehand.
**FZ**  Similarly, the 8 traits with very low variation (160, ..., 231) are hard to exclude before.
**FZ**  Warnings are ok: indicating cases where variance components are set to zero during likelihood optimization.

ERROR : 84 Optimizer (optim) did not achieve convergence (convergence = 10). 
ERROR : 158 Optimizer (optim) did not achieve convergence (convergence = 10). 
ERROR : 160 NA/NaN/Inf in 'y' 
ERROR : 161 NA/NaN/Inf in 'y' 
ERROR : 162 NA/NaN/Inf in 'y' 
ERROR : 163 NA/NaN/Inf in 'y' 
ERROR : 165 NA/NaN/Inf in 'y' 
ERROR : 166 NA/NaN/Inf in 'y' 
ERROR : 168 NA/NaN/Inf in 'y'
ERROR : 231 NA/NaN/Inf in 'y'

Also a bunch of warnings, are any of these a concern?

> warnings()
Warning messages:
1: In metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR,  ... :
  Rows with NAs omitted from model fitting.
...
13: In metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR,  ... :
  There are outcomes with non-positive sampling variances.
14: In metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR,  ... :
  'V' appears to be not positive definite.
16: In metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR,  ... :
  Single-level factor(s) found in 'random' argument. Corresponding 'sigma2' value(s) fixed to 0.
...
50: In metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR,  ... :
  'V' appears to be not positive definite

------
Now that we have a "results" table with each of the meta-analytic means for all effect sizes of interest, we can use this table as part of the Shiny App, which will then be able to back calculate the percentage differences between males and females for mean, variance and coefficient of variance. We'll export and use this in the Shiny App. **Note that I have not dealt with convergence issues in some of these models, and so, this will need to be done down the road**

(Note Susi 31/7/2019: This dataset contains dublicated values, plus no info on what the "traits" mean. I will change Dan N's to one further belwo, that have been cleaned up already
FILE TO USE: METACOMBO (around line 500))

### Merging datasets & removal of non-converged traits

@@

```{r}
results_alltraits_grouping2 <- 
  results_alltraits_grouping %>% 
  left_join(by="id",
             data %>% select(id, parameter_group, procedure = procedure_name, parameter_name) %>%
              # REVIEW -- this bext line replciates precious code, but removes some procedures, is that right?
		  # FZ: No. This creates the data from the complete data set ('data'), which was not carried over into 'results_alltraits_grouping'
		  # FZ: We filter duplicated id's to get only one unique row per id (and there is one id per parameter_name)
              filter(!duplicated(id))
            ) %>%
  # FZ: Below we add 'procedure' (from the previously loaded 'procedures.csv') as a variable
  left_join(by="procedure", 
            procedures %>% distinct()
            )

n <- length(unique(results_alltraits_grouping2$parameter_name)) # 232
```

@@

**Review**: How did we determine this the ones with problems? 
**FZ**: In addition to the sentence one line above, to do: add information about cause of exclusion ( 1. non-convergence of metafor models, or 2. no variation/nonsensical); reference traits to be excluded by trait name (not id number)
**FZ**: excluded because of non-convergence: 84, 158
**FZ**: excluded because no variation/nonsensical: 144, 158, 160,... 
```{r}

#FZ delete# has_problems <- c(84, 144, 158, 160, 161, 162, 163, 165, 166, 167, 168, 221, 222, 231)
#FZ delete# meta_clean <- results_alltraits_grouping2 %>% filter(!id %in% has_problems)
#FZ delete# meta_problem <- results_alltraits_grouping2 %>% filter(id %in% has_problems)

#**FZ**:exclude parameter names for which metafor models didn't converge ("dp t cells", "mzb (cd21/35 high)"), and of parameters that don't harbour enough variation
meta_clean <- results_alltraits_grouping2 %>% 
	  filter(!parameter_name %in% c("dp t cells", "mzb (cd21/35 high)", "number of caudal vertebrae", 
	  "number of cervical vertebrae", "number of digits", "number of lumbar vertebrae", "number of pelvic vertebrae", "number of ribs left",                       
        "number of ribs right", "number of signals", "number of thoracic vertebrae", "total number of acquired events in panel a",
        "total number of acquired events in panel b", "whole arena permanence"))

```
```

**Reveiw**: check against old script -- identical, remove once fixed

```{r, eval=FALSE}
meta_clean.test <- readRDS("../meta_clean.test.rds")
all.equal(meta_clean, meta_clean.test)
all.equal(meta_clean %>% select(-parameter_group), meta_clean.test %>% mutate(id=as.integer(id), GroupingTerm = as.character(GroupingTerm)))
```

@@

meta1 <- meta_clean
length(unique(meta1$procedure)) #18
length(unique(meta1$GroupingTerm)) #9
# Review: next line now 155 not 148 #FZ comment: still 148 for me.
length(unique(meta1$parameter_group)) # 148 levels. To be used as grouping factor for meta-meta analysis / collapsing down based on things that are classified identically in "parameter_group" but have different "parameter_name"
length(unique(meta1$parameter_name)) #218
```

@@

kable(cbind(meta1 %>% count(parameter_group))) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

```{r}
meta1_sub <- meta1 %>%
  # add summary of number of parameter names in each parameter group
  group_by(parameter_group) %>%
  mutate(par_group_size = length(unique(parameter_name)), 
         sampleSize = as.numeric(sampleSize)) %>% 
  ungroup() %>% 
  # Create subsets with > 1 count (par_group_size > 1)
  filter(par_group_size > 1) # 90 observations
```

meta1b <-
  meta1 %>%
  group_by(parameter_group) %>% 
  summarize(par_group_size = length(unique(parameter_name, na.rm = TRUE)))
#this gives a summary of number of parameter names in each parameter group, now it neeeds to get merged it back together

@@

meta1_sub<-subset(meta1,par_group_size >1) # 90 observations   
meta1_sub$sampleSize <- as.numeric(meta1_sub$sampleSize)

# nesting
n_count <- meta1_sub %>%
  group_by(parameter_group) %>%
  mutate(raw_N = sum(sampleSize)) %>%
  nest() %>%
  ungroup()

@@

```{r}
metacombo_final <- metacombo %>%
  group_by(GroupingTerm) %>%
#FZ comment: changed 'nest' to 'nest_legacy' to keep old syntax/functionality
  nest_legacy()

@@

# For all grouping terms
metacombo_final_all <- metacombo %>%
#FZ comment: changed 'nest' to 'nest_legacy' to keep old syntax/functionality
  nest_legacy()

@@

##FZ: 11-12-2019 to here!#

### 2020-04-20 Prep & old/SydneySexDiff WITH comments Felix & Dan.Rmd
Add files via upload

## Load packages & functions

```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  tidy = TRUE
)

@@

1) "Population statistics"

@@

2) Extraction of effect sizes and sample variances

@@

Having loaded the necessary functions, we can get started on the dataset.

We here provide the cleaned dataset, which we have saved in a folder called `export`, as easy starting point. However, the full dataset can be loaded and cleaned using the data cleaning function below. We created this as it much smaller than the .csv.

@@

This step we have already done and provide a cleaned up file which is less computing intensive. However, the cvs is provided in case this is preferred to be attempted, following the steps below:

@@

    # add grouping variable: these were decided based on functional groups and procedures 
    # **Review** - retaining match here to retain ordering of data used in rpevious version
    mutate(parameter_group = group$parameter_group[match(parameter_name, group$parameter_name)] ) %>%
    # usually i would use left_join
        # left_join(by="parameter_name", 
    #   read_csv(here("data", "ParameterGrouping.csv")) %>%
    #     select(-id)
    # ) %>%

@@

```{r}
length(unique(data$parameter_name)) # 232 traits
length(unique(data$parameter_group)) # 161 parameter groups
length(unique(data$procedure_name)) # 26 procedure groups
length(unique(data$biological_sample_id)) # 27147 individial mice   #FZ added 11-12-2019
#number of males and females per strain per production center #FZ added 11-12-2019
sampleSizes <-data %>% group_by(production_center, strain_name) %>% count(biological_sample_id, sex) %>% count(sex) %>% print(n = Inf)
sampleSizes <- write.csv(sampleSizes, file = "export/sampleSizes.csv")
length(unique(sampleSizes$strain_name)) #9
length(unique(data$strain_name)) #9
```

@@

for (t in 160:160) {
  tryCatch(
    {
      results <- data %>% 
        data_subset_parameterid_individual_by_age(t) %>%
        calculate_population_stats() %>%
        create_meta_analysis_effect_sizes()

@@

```{r}
results_alltraits_grouping2 <- 
  results_alltraits_grouping %>% 
  left_join(by="id",
             data %>% select(id, parameter_group, procedure = procedure_name, parameter_name) %>%
              # REVIEW -- this bext line replciates precious code, but removes some procedures, is that right?
		  # FZ: No. This creates the data from the complete data set ('data'), which was not carried over into 'results_alltraits_grouping'
		  # FZ: We filter duplicated id's to get only one unique row per id (and there is one id per parameter_name)
              filter(!duplicated(id))
            ) %>%
  # FZ: Below we add 'procedure' (from the previously loaded 'procedures.csv') as a variable
  left_join(by="procedure", 
            procedures %>% distinct()
            )

@@

#**FZ**:exclude parameter names for which metafor models didn't converge ("dp t cells", "mzb (cd21/35 high)"), and of parameters that don't harbour enough variation
meta_clean <- results_alltraits_grouping2 %>% 
	  filter(!parameter_name %in% c("dp t cells", "mzb (cd21/35 high)", "number of caudal vertebrae", 
	  "number of cervical vertebrae", "number of digits", "number of lumbar vertebrae", "number of pelvic vertebrae", "number of ribs left",                       
        "number of ribs right", "number of signals", "number of thoracic vertebrae", "total number of acquired events in panel a",
        "total number of acquired events in panel b", "whole arena permanence"))

@@

```{r, eval=FALSE}
meta_clean.test <- readRDS("../meta_clean.test.rds")
all.equal(meta_clean, meta_clean.test)
all.equal(meta_clean %>% select(-parameter_group), meta_clean.test %>% mutate(id=as.integer(id), GroupingTerm = as.character(GroupingTerm)))
```

@@

meta1_sub<-subset(meta1,par_group_size >1) # 90 observations   
meta1_sub$sampleSize <- as.numeric(meta1_sub$sampleSize)

@@

```{r}
# Step1
combinedmeta <- bind_rows(robu_all, meta_all)
# glimpse(combinedmeta)

# Steps 2&3
metacombo <- combinedmeta
metacombo$counts <- meta1$par_group_size[match(metacombo$parameter_group, meta1$parameter_group)]
metacombo$procedure2 <- meta1$procedure[match(metacombo$parameter_group, meta1$parameter_group)]
metacombo$GroupingTerm2 <- meta1$GroupingTerm[match(metacombo$parameter_group, meta1$parameter_group)]

@@

 Clean-up and rename

```{r}
metacombo <- metacombo[, c(1, 21:23, 2:13)]
names(metacombo)[3] <- "procedure"
names(metacombo)[4] <- "GroupingTerm"

@@

compare <- metacombo %>%
  group_by(GroupingTerm) %>%
  dplyr::summarize(MeanCVR = mean(lnCVR), MeanVR = mean(lnVR), MeanRR = mean(lnRR))

# compare
```

@@

Re-structure data for each grouping term; delete unused variables

@@

overall2 <- bind_rows(Behaviour, Morphology, Metabolism, Physiology, Immunology, Hematology, Heart, Hearing, Eye, All)
```

@@

##FZ: 11-12-2019 to here!#

@@

malebias_Fig2_alltraits <-
  ggplot(meta.plot2.all.e) +
  aes(x = GroupingTerm, y = percent, fill = sex) +
  geom_col() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "gray40") +
  geom_text(
    data = subset(meta.plot2.all.e, samplesize != 0), aes(label = samplesize), position = position_stack(vjust = .5),
    color = "white", size = 3.5
  ) +
  facet_grid(
    cols = vars(trait), rows = vars(label), labeller = label_wrap_gen(width = 18),
    scales = "free", space = "free"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_bw(base_size = 18) +
  theme(
    strip.text.y = element_text(angle = 270, size = 10, margin = margin(t = 15, r = 15, b = 15, l = 15)),
    strip.text.x = element_text(size = 12),
    strip.background = element_rect(colour = NULL, linetype = "blank", fill = "gray90"),
    text = element_text(size = 14),
    panel.spacing = unit(0.5, "lines"),
    panel.border = element_blank(),
    axis.line = element_line(),
    panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
    panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  ) +
  coord_flip()

>>>>

## Loading packages & custom functions

```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  tidy = TRUE
)
```

@@

1) "Population statistics": "calculate_population_stats"
This function groups animals from the same strain and same insitiution together. This is done for each trait seoarately, and only for traits that have been measured in both sexes. Any group containing fewer than 5 individuals is excluded.

@@

2) Extraction of effect sizes and sample variances: "create_meta_analysis_effect_sizes"

@@

@@

This step we have already done and provide a cleaned up file which is less computing intensive and which we have saved in a folder called `export`. However, the cvs is provided in case this is preferred to be attempted, following the steps below:

@@

    # add grouping variable: these were decided based on functional groups and procedures 
    mutate(parameter_group = group$parameter[match(parameter_name, group$parameter_name)] ) %>%
    # usually i would use left_join
        # left_join(by="parameter_name", 
    #   read_csv(here("data", "ParameterGrouping.csv")) %>%
    #     select(-id)
    # ) %>%

@@

```{r}
length(unique(data$parameter_name)) # 232 traits
length(unique(data$parameter_group)) # 161 parameter groups
length(unique(data$procedure_name)) # 26 procedure groups
length(unique(data$biological_sample_id)) # 27147 individial mice   #FZ added 11-12-2019
#number of males and females per strain per production center #FZ added 11-12-2019  # FOR TABLE SAMPLE SIZES
data %>% group_by(production_center, strain_name) %>% count(biological_sample_id, sex) %>% count(sex) %>% print(n = Inf)

@@

for (t in 1:n) {
  tryCatch(
    {
      results <- data %>% 
        data_subset_parameterid_individual_by_age(t) %>%
        calculate_population_stats() %>%
        create_meta_analysis_effect_sizes()

@@

```{r}
results_alltraits_grouping2 <- 
  results_alltraits_grouping %>% 
  left_join(by="id",
             data %>% select(id, parameter_group, procedure = procedure_name, procedure_name, parameter_name) %>% #susi: check parateer group
              # REVIEW -- this bext line replciates precious code, but removes some procedures, is that right?
		  # FZ: No. This creates the data from the complete data set ('data'), which was not carried over into 'results_alltraits_grouping'
		  # FZ: We filter duplicated id's to get only one unique row per id (and there is one id per parameter_name)
              filter(!duplicated(id))
            ) %>%
  # FZ: Below we add 'procedure' (from the previously loaded 'procedures.csv') as a variable
  left_join(by="procedure", 
            procedures %>% distinct()
            )

@@

#**FZ**: We exclude 14 parameter names for which metafor models didn't converge ("dp t cells", "mzb (cd21/35 high)"), and of parameters that don't harbour enough variation
meta_clean <- results_alltraits_grouping2 %>% 
	  filter(!parameter_name %in% c("dp t cells", "mzb (cd21/35 high)", "number of caudal vertebrae", 
	  "number of cervical vertebrae", "number of digits", "number of lumbar vertebrae", "number of pelvic vertebrae", "number of ribs left",                       
        "number of ribs right", "number of signals", "number of thoracic vertebrae", "total number of acquired events in panel a",
        "total number of acquired events in panel b", "whole arena permanence"))

@@

```{r, eval=FALSE}
meta_clean.test <- readRDS("../meta_clean.test.rds")  #FZ comment: doesn't work for me. '../' is for the github location?
all.equal(meta_clean, meta_clean.test)
all.equal(meta_clean %>% select(-parameter_group), meta_clean.test %>% mutate(id=as.integer(id), GroupingTerm = as.character(GroupingTerm)))
```

@@

meta1_sub <- subset(meta1,par_group_size >1) # 90 observations   
meta1_sub$sampleSize <- as.numeric(meta1_sub$sampleSize)

@@

```{r}
# Step1 (columns are matched by name (in our case, 'parameter_group'), and any missing columns will be filled with NA) #FZ added explanation of bibd_rows
combinedmeta <- bind_rows(robu_all, meta_all)
# glimpse(combinedmeta)

# Steps 2&3 (add information about number of traits in a parameter group, procedure, and grouping term) #FZ added expalnation in brackets
metacombo <- combinedmeta
metacombo$counts <- meta1$par_group_size[match(metacombo$parameter_group, meta1$parameter_group)] 
metacombo$procedure2 <- meta1$procedure[match(metacombo$parameter_group, meta1$parameter_group)]
metacombo$GroupingTerm2 <- meta1$GroupingTerm[match(metacombo$parameter_group, meta1$parameter_group)]

@@

Clean-up, reorder, and rename

```{r}
metacombo <- metacombo[, c(1, 21:23, 2:13)] # SUSI CHANGE TO NAMES!!!
names(metacombo)[3] <- "procedure" #change "3" to name
names(metacombo)[4] <- "GroupingTerm" #change "4" to name

@@

metacombo %>%
  group_by(GroupingTerm) %>%
  dplyr::summarize(MeanCVR = mean(lnCVR), MeanVR = mean(lnVR), MeanRR = mean(lnRR))

@@

Re-structure data for each grouping term; delete unused variables  #FZ comment: referencing of cells below doesn't depend on previous oreding of the data, i.e. only changes if output structure from metafor::rma.uni changes

@@

overall2 <- bind_rows(Behaviour, Morphology, Metabolism, Physiology, Immunology, Hematology, Heart, Hearing, Eye, All) #FZ: warnings are ok
```

@@

@@

# add summary row ('All') and re-arrange rows into correct order for plotting #FZ added

meta.plot2.all.f <- meta.plot2.all.e %>% group_by(trait, sex) %>% 
	summarise(GroupingTerm = "All", malebias = sum(malebias), femalebias = sum(femalebias), total = malebias + femalebias, 
	label = "All traits", samplesize = sum(samplesize)) %>%
	mutate(percent = ifelse(sex == "femalepercent", femalebias*100/(malebias+femalebias), malebias*100/(malebias+femalebias))) %>%
	bind_rows(meta.plot2.all.e, .) %>%
	mutate(rownumber = row_number()) %>%
	.[c(37, 1:9, 39, 10:18, 38, 19:27, 40, 28:36), ]

meta.plot2.all.f$GroupingTerm <- factor(meta.plot2.all.f$GroupingTerm, levels = c("Behaviour", "Morphology", "Metabolism", "Physiology", "Immunology", "Hematology", "Heart", "Hearing", "Eye", "All")) 
meta.plot2.all.f$GroupingTerm <- factor(meta.plot2.all.f$GroupingTerm, rev(levels(meta.plot2.all.f$GroupingTerm)))

malebias_Fig2_alltraits <-
  ggplot(meta.plot2.all.f) +
  aes(x = GroupingTerm, y = percent, fill = sex) +
  geom_col() +
  geom_hline(yintercept = 50, linetype = "dashed", color = "gray40") +
  geom_text(
    data = subset(meta.plot2.all.f, samplesize != 0), aes(label = samplesize), position = position_stack(vjust = .5),
    color = "white", size = 3.5
  ) +
  facet_grid(
    cols = vars(trait), rows = vars(label), labeller = label_wrap_gen(width = 18),
    scales = "free", space = "free"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_bw(base_size = 18) +
  theme(
    strip.text.y = element_text(angle = 270, size = 10, margin = margin(t = 15, r = 15, b = 15, l = 15)),
    strip.text.x = element_text(size = 12),
    strip.background = element_rect(colour = NULL, linetype = "blank", fill = "gray90"),
    text = element_text(size = 14),
    panel.spacing = unit(0.5, "lines"),
    panel.border = element_blank(),
    axis.line = element_line(),
    panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
    panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  ) +
  coord_flip()

### 2020-04-20 Prep & old/SydneySexDiffFelix&Dan.Rmd
Add files via upload

## Loading packages & custom functions

```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  tidy = TRUE
)
```

@@

1) "Population statistics": "calculate_population_stats"
This function groups animals from the same strain and same insitiution together. This is done for each trait seoarately, and only for traits that have been measured in both sexes. Any group containing fewer than 5 individuals is excluded.

@@

2) Extraction of effect sizes and sample variances: "create_meta_analysis_effect_sizes"

@@

@@

This step we have already done and provide a cleaned up file which is less computing intensive and which we have saved in a folder called `export`. However, the cvs is provided in case this is preferred to be attempted, following the steps below:

@@

    # add grouping variable: these were decided based on functional groups and procedures 
    mutate(parameter_group = group$parameter[match(parameter_name, group$parameter_name)] ) %>%
    # usually i would use left_join
        # left_join(by="parameter_name", 
    #   read_csv(here("data", "ParameterGrouping.csv")) %>%
    #     select(-id)
    # ) %>%

>>>>

## Load packages & functions

```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  tidy = TRUE
)

@@

1) "Population statistics"

@@

2) Extraction of effect sizes and sample variances

@@

Having loaded the necessary functions, we can get started on the dataset.

We here provide the cleaned dataset, which we have saved in a folder called `export`, as easy starting point. However, the full dataset can be loaded and cleaned using the data cleaning function below. We created this as it much smaller than the .csv.

@@

This step we have already done and provide a cleaned up file which is less computing intensive. However, the cvs is provided in case this is preferred to be attempted, following the steps below:

@@

    # add grouping variable: these were decided based on functional groups and procedures 
    # **Review** - retaining match here to retain ordering of data used in rpevious version
    mutate(parameter_group = group$parameter[match(parameter_name, group$parameter_name)] ) %>%
    # usually i would use left_join
        # left_join(by="parameter_name", 
    #   read_csv(here("data", "ParameterGrouping.csv")) %>%
    #     select(-id)
    # ) %>%
